---
title: "Quantium Virtual Internship - Retail Strategy and Analytics - Task 1"
output: html_notebook
---


```{r  setup, include=FALSE}
# set options for R markdown knitting
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(linewidth=80)
```


```{r  knitr line wrap setup, include=FALSE}
# set up line wrapping in MD knit output
library(knitr)
hook_output = knit_hooks$get("output")
knit_hooks$set(output = function(x, options)
{
 # this hook is used only when the linewidth option is not NULL
 if (!is.null(n <- options$linewidth))
 {
 x = knitr:::split_lines(x)
 # any lines wider than n should be wrapped
 if (any(nchar(x) > n))
 x = strwrap(x, width = n)
 x = paste(x, collapse = "\n")
 }
 hook_output(x, options)
})

```

## Load required libraries and datasets

```{r 0 Load Libraries, results = 'hide'}

#### Example code to install packages
#install.packages("data.table")
#### Load required libraries
library(data.table)
library(ggplot2)
library(ggmosaic)
library(readr)
library(tm)
library(wordcloud)
library(dplyr)
library(stringr)
library(ggplot2)
library(gganimate)
library(tidyverse)
#library(qdap)

```
#### Point the filePath to where you have downloaded the datasets to and
#### assign the data files to data.tables

```{r}
filePath <- "/Volumes/Seagate_Haykay/Quantium_Data_Analytics"
setwd(filePath)
```
```{r}
getwd()
```




```{r}

filePath <- "/Volumes/Seagate_Haykay/Quantium_Data_Analytics"
transactionData <- fread(paste0(filePath,"/QVI_transaction_data.csv"))
customerData <- fread(paste0(filePath,"/QVI_purchase_behaviour.csv"))
```


## Exploratory data analysis
The first step in any analysis is to first understand the data. Let's take a look
at each of the datasets provided.
### Examining transaction data
We can use `str()` to look at the format of each column and see a sample of the
data. As we have read in the dataset as a `data.table` object, we can also run
`transactionData` in the console to see a sample of the data or use
`head(transactionData)` to look at the first 10 rows.


```{r }
head(transactionData)
```



```{r Examining transaction data}
#### Examine customer data
head(customerData)
```


Let's check if columns we would expect to be numeric are in numeric form and date
columns are in date format.


```{r Convert DATE to date format}
#### Convert DATE column to a date format
#### A quick search online tells us that CSV and Excel integer dates begin on 30 Dec 1899
transactionData$DATE <- as.Date(transactionData$DATE, origin = "1899-12-30")


transactionData$TOT_SALES <- str_replace(transactionData$TOT_SALES, ",", ".")
transactionData$TOT_SALES <- as.double(transactionData$TOT_SALES)
head(transactionData)
```


We should check that we are looking at the right products by examining PROD_NAME.



```{r Summary of PROD_NAME}
#### Examine PROD_NAME
summary(transactionData$PROD_NAME)

PROD_NAME <- (transactionData$PROD_NAME)

head(PROD_NAME)
```


Looks like we are definitely looking at potato chips but how can we check that
these are all chips? We can do some basic text analysis by summarising the
individual words in the product name.


```{r Further examine PROD_NAME}
#### Examine the words in PROD_NAME to see if there are any incorrect entries
#### such as products that are not chips
productWords <- data.table(unlist(strsplit(unique(transactionData[, transactionData$PROD_NAME]), "
")))
setnames(productWords, 'words')
head(transactionData)
```


As we are only interested in words that will tell us if the product is chips or
not, let's remove all words with digits and special characters such as '&' from our
set of product words. We can do this using `grepl()`.



```{r}
# Remove digits, and special characters, and then sort the distinct words by frequency of occurrence.

#### Removing digits
transactionData$PROD_NAME_NO_CHAR <- data.table(gsub("\\w*[0-9]+\\w*", "", transactionData$PROD_NAME))

head(transactionData)
```



```{r}
#### Removing special characters
# as.data.frame(gsub("[[:punct:]]", "", as.matrix(df))) 
transactionData$PROD_NAME_NO_CHAR <- as.data.frame(gsub("[[:punct:]]", "", as.matrix(transactionData$PROD_NAME_NO_CHAR)))
# setnames(No.specialchar.digits, 'PROD_NAME_NO_CHAR')
head(transactionData)
```


collapse column into one giant string


```{r}
review_text <- paste(transactionData$PROD_NAME_NO_CHAR, collapse=" ")
```



```{r}
# review_text
```



```{r}
# setting up source and corpus
review_source <- VectorSource(review_text)
corpus <- Corpus(review_source)
```



```{r}
# Cleaning
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, removeWords, stopwords(kind = "en"))
```


```{r}
# Making a document-term matrix
dtm <- DocumentTermMatrix(corpus)
#dtm2 <- as.matrix(dtm)
m <- as.matrix(dtm)
v <- sort(colSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=500, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```


```{r}
#### sorting them by this frequency in order of highest to lowest frequency
#### Let's look at the most common words by counting the number of times a word appears and
# finding the most frequent term
frequency <- colSums(m)
freq <- sort(frequency, decreasing = TRUE)
head(freq, n = 20L)
```


There are salsa products in the dataset but we are only interested in the chips
category, so let's remove these.



```{r}
#### Remove salsa products
transactionData[, SALSA := grepl("salsa", tolower(PROD_NAME_NO_CHAR))]
transactionData <- transactionData[SALSA == FALSE, ][, SALSA := NULL]
head(transactionData)
```



```{r}
review_text <- paste(transactionData$PROD_NAME_NO_CHAR, collapse=" ")

review_source <- VectorSource(review_text)
corpus <- Corpus(review_source)

corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, removeWords, stopwords(kind = "en"))

dtm <- DocumentTermMatrix(corpus)
dtm2 <- as.matrix(dtm)

frequency <- colSums(dtm2)
freq <- sort(frequency, decreasing = TRUE)
#head(freq, n = 20L)
```


```{r}

head(freq, n = 20L)
```



```{r initial summary}
#### Summarise the data to check for nulls and possible outliers
summary(transactionData)
```

There are no nulls in the columns but product quantity appears to have an outlier
which we should investigate further. Let's investigate further the case where 200
packets of chips are bought in one transaction.


```{r}
#### Filter the dataset to find the outlier
#library(ggplot2)

ggplot(transactionData) +
  aes(x ="", y = PROD_QTY) +
  geom_boxplot(fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
# library(dplyr)
outliers <- transactionData %>% filter(PROD_QTY >= 199)
outliers
```



```{r}
#### Let's see if the customer has had other transactions
customer_transactions <- transactionData %>% filter(LYLTY_CARD_NBR == 226000)
customer_transactions
```


```{r}
#### Filter out the customer based on the loyalty card number
# remove rows in r - subset function with multiple conditions
transactionData <- transactionData[!(transactionData$LYLTY_CARD_NBR==226000),]

#df2<-df1[!(df1$Name=="George" | df1$Name=="Andrea"),]

#df2

# transactionData$LYLTY_CARD_NBR
```


```{r}
#### Re-examine transaction data
summary(transactionData)
```



```{r}
#### Count the number of transactions by date

length(unique(transactionData[["DATE"]]))
```


There's only 364 rows, meaning only 364 dates which indicates a missing date. Let's
create a sequence of dates from 1 Jul 2018 to 30 Jun 2019 and use this to create a
chart of number of transactions over time to find the missing date.


```{r fig.align = "center"}

#### Create a sequence of dates and join this the count of transactions by date
# Over to you - create a column of dates that includes every day from 1 Jul 2018 to 30 Jun 2019, and join it onto the data to fill in the missing day.
transactions_by_day <-
        data.table(table(as.Date(transactionData$DATE)))
        setnames(transactions_by_day, c('DATE', 'N'), skip_absent=TRUE)
        
transactions_by_day_2 <- transactionData[, .N, DATE][order(DATE)]
transactions_by_day_2

transactions_by_day$DATE <- as.Date(transactions_by_day$DATE, origin = "1899-12-30")
print(transactions_by_day)
```



```{r fig.align = "center"}
#### Setting plot themes to format graphs
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
#### Plot transactions over time
ggplot(transactions_by_day, aes(x = DATE, y = N)) +
geom_line() +
labs(x = "Day", y = "Number of transactions", title = "Transactions over time") +
scale_x_date(breaks = "1 month") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```


```{r fig.align = "center"}
#### Filter to December and look at individual days
december <- transactions_by_day %>% select(DATE, N) %>% filter(DATE >= as.Date("2018-12-01") & DATE <= as.Date("2018-12-31"))
december
```


```{r fig.align = "center"}
#### Setting plot themes to format graphs
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
#### Plot transactions over time
ggplot(december, aes(x = DATE, y = N)) +
geom_point() +
labs(x = "Day", y = "Number of transactions", title = "Transactions over time") +
scale_x_date(breaks = "1 day") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```


We can see that the increase in sales occurs in the lead-up to Christmas and that
there are zero sales on Christmas day itself. This is due to shops being closed on
Christmas day.
Now that we are satisfied that the data no longer has outliers, we can move on to
creating other features such as brand of chips or pack size from PROD_NAME. We will
start with pack size.


```{r Create pack size}
#### Pack size
#### We can work this out by taking the digits that are in PROD_NAME
filePath <- "/Volumes/Seagate_Haykay/Quantium_Data_Analytics"
transactionData2 <- fread(paste0(filePath,"/QVI_transaction_data.csv"))
transactionData2[, PACK_SIZE := parse_number(PROD_NAME)]



#### Always check your output
#### Let's check if the pack sizes look sensible
transactionData2[, .N, PACK_SIZE][order(N)]
```
The largest size is 380g and the smallest size is 70g - seems sensible!

```{r }
#### Let's plot a histogram of PACK_SIZE since we know that it is a categorical variable and not a continuous variable even though it is numeric.
# Plot a histogram showing the number of transactions by pack size.

PACK_SIZE <- transactionData2$PACK_SIZE
h <- hist(PACK_SIZE,
main="Number of Transaction by Pack Size",
xlab="PACK SIZE",
xlim=c(60,390),
ylim=c(1400,120000),
breaks=10,
col="skyblue",
# freq=FALSE
)
text(h$mids,h$counts,labels=h$counts, adj=c(0.5, -0.5))
```

Now to create brands, we can use the first word in PROD_NAME to work out the brand
name...
```{r Create brand name}
#### Brands
# Over to you! Create a column which contains the brand of the product, by extracting it from the product name.
# library(stringr)
transactionData2$BRAND <- str_extract(transactionData2$PROD_NAME,"(\\w+)") 
head(transactionData2, 20L)
```



```{r}
transactionData2[, .N, BRAND][order(BRAND)]

```
Some of the brand names look like they are of the same brands - such as RED and
RRD, which are both Red Rock Deli chips. Let's combine these together.

```{r Clean brand names}
#### Clean brand names
transactionData2[BRAND == "Red", BRAND := "RRD"]
transactionData2[BRAND == "Infzns", BRAND := "Infuzions"]
transactionData2[BRAND == "Dorito", BRAND := "Doritos"]
transactionData2[BRAND == "Natural", BRAND := "NCC"]
transactionData2[BRAND == "Smith", BRAND := "Smiths"]
transactionData2[BRAND == "Snbts", BRAND := "Sunbites"]
transactionData2[BRAND == "Grain", BRAND := "GrnWves"]
transactionData2[BRAND == "WW", BRAND := "Woolworths"]
transactionData2[, .N, BRAND][order(BRAND)]

```
### Examining customer data
Now that we are happy with the transaction dataset, let's have a look at the
customer dataset.

```{r 1 Exploratory data analysis}
#### Examining customer data
summary(customerData)
```


```{r}
# distributions of LYLTY_CARD_NBR key columns
# library(ggplot2)
# library(gganimate)
# library(tidyverse)

ggplot(customerData, aes(x=as.factor(PREMIUM_CUSTOMER), fill=as.factor(PREMIUM_CUSTOMER) )) + 
  geom_bar( ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position="none") +
  coord_flip() +
  ggtitle("Customers")
```

```{r}
ggplot(customerData, aes(x=as.factor(LIFESTAGE), fill=as.factor(LIFESTAGE) )) + 
  geom_bar( ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position="none") +
  coord_flip() +
  ggtitle("LIFESTAGE")
```
As the number of rows in `data` is the same as that of `transactionData`, we can be
sure that no duplicates were created. This is because we created `data` by setting
`all.x = TRUE` (in other words, a left join) which means take all the rows in
`transactionData` and find rows with matching values in shared columns and then
joining the details in these rows to the `x` or the first mentioned table.

```{r}
#### Merge transaction data to customer data
transactionData2$DATE <- as.Date(transactionData2$DATE, origin = "1899-12-30")
data <- merge(transactionData2, customerData, all.x = TRUE)
data
# total <- merge(transactionData2, customerData,by="LYLTY_CARD_NBR")
```
Let's also check if some customers were not matched on by checking for nulls.
```{r Check for missing customer details}
# summary(data)

sum(is.na(data))

# Find missing values in a column of a data frame

# which(is.na(expl_data1$x1))


# In order to get the positions of each column in your data set, you can use the apply() function

# apply(is.na(expl_data1), 2, which) 

```

Note that if you are continuing with Task 2, you may want to retain this dataset which you can write out as a csv

```{r Code to save dataset as a csv}
filePath <- "/Volumes/Seagate_Haykay/Quantium_Data_Analytics"
fwrite(data, paste0(filePath,"QVI_data.csv"))
write_csv(data,"/Volumes/Seagate_Haykay/Quantium_Data_Analytics/QVI_data.csv")
```


## Data analysis on customer segments
Now that the data is ready for analysis, we can define some metrics of interest to
the client:
- Who spends the most on chips (total sales), describing customers by lifestage and
how premium their general purchasing behaviour is
- How many customers are in each segment
- How many chips are bought per customer by segment
- What's the average chip price by customer segment
We could also ask our data team for more information. Examples are:
- The customer's total spend over the period and total spend for each transaction
to understand what proportion of their grocery spend is on chips
- Proportion of customers in each customer segment overall to compare against the
mix of customers who purchase chips


```{r fig.width = 10, fig.align = "center"}
# library(dplyr)
#### Total sales by LIFESTAGE and PREMIUM_CUSTOMER
library(reshape2)
        
total_sales_lifestage_customers <- select(data, TOT_SALES, LIFESTAGE, PREMIUM_CUSTOMER)

total_sales_lifestage_customers$R_TOT_SALES <- str_replace(total_sales_lifestage_customers$TOT_SALES,",",".")

total_sales_lifestage_customers$R_TOT_SALES <- as.double(total_sales_lifestage_customers$R_TOT_SALES)
total_sales_lifestage_customers$TOT_SALES <- NULL
head(total_sales_lifestage_customers)


total_sales_plot <- total_sales_lifestage_customers %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(sales = sum(R_TOT_SALES)) %>%
  arrange(.,  sales)

total_sales_plot
```


```{r fig.width = 10, fig.align = "center"}}
# bar plot, fill to discrete column, color to static value
ggplot(total_sales_plot, aes(LIFESTAGE, sales, fill = PREMIUM_CUSTOMER)) + geom_col(position = "dodge") + coord_flip()
```

```{r fig.width = 10, fig.align = "center"}}
#### Number of customers by LIFESTAGE and PREMIUM_CUSTOMER
# Calculate the summary of number of customers by those dimensions and create a plot.
no_of_customers_lifestage_customers <- select(data, LIFESTAGE, PREMIUM_CUSTOMER)



total_customers <- no_of_customers_lifestage_customers %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(Count = n()) %>%
  arrange(., Count)

ggplot(total_customers, aes(LIFESTAGE, Count, fill = PREMIUM_CUSTOMER)) + 
  geom_col(position = "dodge") +
  labs(title = "No. of Customers by LIFESTAGE and PREMIUM_CUSTOMER") +
  ylim(0, 30000) +
  scale_y_continuous(breaks = seq(from = 0, to = 30000, by = 5000)) +
  coord_flip()
```
There are more Mainstream - young singles/couples and Mainstream - retirees who buy
chips. This contributes to there being more sales to these customer segments but
this is not a major driver for the Budget - Older families segment.

Higher sales may also be driven by more units of chips being bought per customer.
Let's have a look at this next.

```{r fig.width = 10, fig.align = "center"}
#### Average number of units per customer by LIFESTAGE and PREMIUM_CUSTOMER

# Calculate and plot the average number of units per customer by those two dimensions.
average_sales_lifestage_customers <- select(data, TOT_SALES, LIFESTAGE, PREMIUM_CUSTOMER)
average_sales_lifestage_customers$TOT_SALES <- str_replace(average_sales_lifestage_customers$TOT_SALES, ",", ".")
average_sales_lifestage_customers$TOT_SALES <- as.double(average_sales_lifestage_customers$TOT_SALES)

average_sales_plot <- average_sales_lifestage_customers %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(ave_sales = mean(TOT_SALES)) %>%
  arrange(.,  ave_sales)
average_sales_plot
```

```{r}
ggplot(average_sales_plot, aes(LIFESTAGE, ave_sales, fill = PREMIUM_CUSTOMER)) + 
  geom_col(position = "dodge") +
  labs(title = "Average number of units per customer by LIFESTAGE and PREMIUM_CUSTOMER") +
  ylim(6, 8) +
  scale_y_continuous(breaks = seq(from = 0, to = 8, by = 1)) +
  coord_flip()
```

```{r fig.width = 10, fig.align = "center"}
#### Average price per unit by LIFESTAGE and PREMIUM_CUSTOMER
# Calculate and plot the average price per unit sold (average sale price) by those two customer dimensions.
average_price_lifestage_customers <- select(data, LIFESTAGE, PREMIUM_CUSTOMER, TOT_SALES, PROD_QTY)

average_price_lifestage_customers$TOT_SALES <- str_replace(average_sales_lifestage_customers$TOT_SALES, ",", ".")

average_price_lifestage_customers$TOT_SALES <- as.double(average_sales_lifestage_customers$TOT_SALES)

average_price_lifestage_customers$AVG_PRICE_PER_UNIT <- with(average_price_lifestage_customers, TOT_SALES / PROD_QTY)


average_price_per_unit <- average_price_lifestage_customers %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(ave_price = mean(AVG_PRICE_PER_UNIT)) %>%
  arrange(.,  ave_price)
average_price_per_unit
#avg_price_plot <- as.data.frame(average_price_lifestage_customers)
# ggplot(average_price_lifestage_customers, aes(LIFESTAGE, AVG_PRICE_PER_UNIT, fill = PREMIUM_CUSTOMER)) + 
#   geom_col(position = "dodge") +
#   labs(title = "Average price per unit by LIFESTAGE and PREMIUM_CUSTOMER") +
#   coord_flip()
```


What do you need to run an independent t-test?
In order to run an independent t-test, you need the following:

One independent, categorical variable that has two levels/groups.
One continuous dependent variable.
```{r}
#### Perform an independent t-test between mainstream vs premium and budget midage and young singles and couples
# Perform a t-test to see if the difference is significant.

# Let's group our data into mainstream and non-mainstream
average_price_per_unit[average_price_per_unit == "Budget" | average_price_per_unit == "Premium"] <- "nonMainstream"

mainstream_compare <- average_price_per_unit %>% filter(LIFESTAGE == 'YOUNG SINGLES/COUPLES' | LIFESTAGE == 'MIDAGE SINGLES/COUPLES', PREMIUM_CUSTOMER == 'Mainstream')
mainstream_compare 


non_mainstream_compare <- average_price_per_unit %>% filter(LIFESTAGE == 'YOUNG SINGLES/COUPLES' | LIFESTAGE == 'MIDAGE SINGLES/COUPLES', PREMIUM_CUSTOMER == 'nonMainstream')

non_mainstream_compare <- non_mainstream_compare %>%
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>%
  summarise(ave_price = mean(ave_price)) %>%
  arrange(.,  ave_price)
non_mainstream_compare

```


```{r}
# p <- ggplot(mainstream_compare , aes(x=ave_price)) + 
#   geom_histogram(binwidth=0.01)
# p
```






```{r}
# mainstream <-  mainstream_compare$ave_price
# mainstream 
# non_mainstream <- non_mainstream_compare$ave_price 
# non_mainstream


compare_plots <- rbind(mainstream_compare, non_mainstream_compare)
compare_plots

# ggplot(compare_plots, aes(PREMIUM_CUSTOMER, fill = ave_price)) +
#    geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity')
```

independent t-test
t.test (dependentVariable ~ independentVariable, var.equal=TRUE, data = dataframe)


```{r}

descriptives = compare_plots %>% group_by(PREMIUM_CUSTOMER) %>% summarise(mean = mean(ave_price), sd = sd(ave_price), n = n())
descriptives

print("===============================================")


mean_difference = descriptives[1,2] - descriptives[2,2]

mean_difference

print("===============================================")
t.test (ave_price ~ PREMIUM_CUSTOMER, var.equal=TRUE, data = compare_plots)
```
The results show that the mean price per unit  in the "mainstream" was 4.04$ (to 2 decimal places) with a standard deviation of 0.05$. There were 2 participants in the mainstream group (Young and midage)

The mean price per unit  in the "non-mainstream" was 3.69$ (to 2 decimal places) with a standard deviation of 0.07$. There were 2 participants in the non-mainstream group (Young and midage).

Therefore, there was a mean difference of 0.34$ between the mainstream and non-mainstream group, with price per unit  value 0.34 higher in the control group (i.e., 4.035793 – 3.691153 = 0.34). The 95% confidence interval (CI) of the mean difference was from 0.08$ (for the lower bound value; to 2 decimal places) to 0.61$ (for the upper bound value).

Independent-samples t-test

From the t.test output above, the obtained t-value (t) is 5.64 (to 2 decimal places) and the degrees of freedom (df) are 2 (i.e., for an independent-samples t-test the degrees of freedom are equal to the sample size minus 2, so in our example, 4 – 2 = 19). The statistical significance (p-value) of the independent-samples t-test is 0.03.

The p-value is used to determine if there is a statistically significant difference in the mean score of the dependent variable between the two independent groups. If p < .05 (i.e., if p is less than .05), there is a statistically significant difference in mean price per unit between the mainstream and non-mainstream group. If p > .05 (i.e., if p is greater than .05), there is not a statistically significant difference in mean price per unit between the mainstream and non-mainstream group. Since p = .03 in our example, there is a statistically significant difference in mean price per unit between the mainstream and non-mainstream group (i.e., there is a mean difference in the population and not only the sample that was studied).




## Deep dive into specific customer segments for insights
We have found quite a few interesting insights that we can dive deeper into.
We might want to target customer segments that contribute the most to sales to
retain them or further increase sales. Let's look at Mainstream - young
singles/couples. For instance, let's find out if they tend to buy a particular
brand of chips.

```{r fig.align = "center"}
#### Deep dive into Mainstream, young singles/couples
# Work out of there are brands that these two customer segments prefer more than others. You could use a technique called affinity analysis or a-priori analysis (or any other method if you prefer)

mainstream_YSC <- data %>% 
  filter(LIFESTAGE == 'YOUNG SINGLES/COUPLES', PREMIUM_CUSTOMER == 'Mainstream')

mainstream_YSC <- select(mainstream_YSC, LYLTY_CARD_NBR, LIFESTAGE, PREMIUM_CUSTOMER, BRAND , TOT_SALES, PROD_QTY, PROD_NAME)

mainstream_YSC$TOT_SALES <- str_replace(mainstream_YSC$TOT_SALES, ",", ".")

mainstream_YSC$TOT_SALES <- as.double(mainstream_YSC$TOT_SALES)
mainstream_YSC_New <- mainstream_YSC


mainstream_YSC_New



# mainstream_YSC_New <- mainstream_YSC_New %>%
#   group_by(BRAND) %>%
#   summarise(tot_sales_by_brand = sum(TOT_SALES)) %>%
#   arrange(.,  tot_sales_by_brand)
# mainstream_YSC_New


```


```{r}

# library(arules)
# library(arulesViz)
# library(IRdisplay)


mainstream_YSC_New
```


```{r}

# mainstream_YSC_New$LIFESTAGE <- NULL
# mainstream_YSC_New$PREMIUM_CUSTOMER <- NULL
# mainstream_YSC_New$TOT_SALES <- NULL
# mainstream_YSC_New$PROD_QTY <- NULL
# mainstream_YSC_New$PROD_NAME <- NULL

head(mainstream_YSC_New)

```



```{r Getting the shopping baskets}
# mainstream_YSC_New_2 <- mainstream_YSC_New %>%
#   group_by(LYLTY_CARD_NBR) %>%
#      mutate(bars_by_foo = paste0(BRAND, collapse = ","))
# 
# mainstream_YSC_New_2$BRAND <- NULL


mainstream_YSC_New_2 <- mainstream_YSC_New %>%
  group_by(LYLTY_CARD_NBR) %>%
     summarise(basket = as.vector(list(BRAND)))
str(mainstream_YSC_New_2)
```





```{r compute transactions}
transactions <- as(mainstream_YSC_New_2$basket, "transactions")
hist(size(transactions), breaks = 0:20)
```


```{r more detailed graph}
par(mfrow=c(1,1))

hist(size(transactions), breaks = 0:20, xaxt="n", ylim = c(0, 4000),
     main = "Number of Purchases per Customer", xlab = "Numboe or Items")
axis(1, at=seq(0,20,by=5), cex.axis=0.7)

mtext(paste("Total:", length(transactions), "customers,", sum(size(transactions)), "brands purchases"))
```



```{r}
basketSizes <- size(transactions)
summary(basketSizes)

quantile(basketSizes, probs = seq(0,1,0.05))
```




```{r get average customer purchase amount}
meanBasketAmt <- aggregate(TOT_SALES ~ LYLTY_CARD_NBR, data=mainstream_YSC_New, sum)
summary(meanBasketAmt)
```

```{r get average price per item}
meanBasketPrice <- aggregate(TOT_SALES ~ LYLTY_CARD_NBR, data=mainstream_YSC_New, mean)
summary(meanBasketPrice)
```


```{r find out which brands customers are buying}
item_frequencies <- itemFrequency(transactions)
sum(item_frequencies)

brandCount <- round((item_frequencies/sum(item_frequencies)) * sum(basketSizes))
summary(brandCount)
```


```{r get top 10 brands in the basket}
orderedProducts <- sort(brandCount, decreasing = TRUE)
orderedProducts[1:10]
```

```{r How many times does the most purchased brand appear in the cart = 38 %}
orderedProducts[1]/dim(transactions)[1]
```


Mining Associations With Apriori

Mine frequent itemsets, association rules or association hyperedges using the Apriori algorithm. The Apriori algorithm employs level-wise search for frequent itemsets. The implementation of Apriori used includes some improvements (e.g., a prefix tree and item sorting).

```{r Plot the top affinity brands}
# Get affinity plot and data using Apriori Algorithms
# support <- 0.1
# support <- 0.05

support <- 0.05

itemsets <- apriori(transactions, parameter = list(target = "frequent itemsets", supp=support, minlen=2), control = list(verbose=FALSE)) 
par(mar=c(5,18,2,2)+.1)

sets_order_supp <- DATAFRAME(sort(itemsets, by="support", decreasing = F))

barplot(sets_order_supp$support, names.arg=sets_order_supp$items, xlim=c(0,0.12), horiz =T, las = 2, cex.names = .6, main= "Frequent Brands")

mtext(paste("support:", support), padj = .8)
```


```{r use apriori algorithm to determine rules - number of affinities, and then output the top 30 to a csv file}
brand_baskets <- transactions[basketSizes]

num <- 10000/dim(brand_baskets)[1] # Number is greater than 1 then use support above 0.02
num
rules <- apriori(brand_baskets, parameter = list(supp=0.05, conf = 0.01, maxlen=3), control = list(verbose=FALSE))
plot(rules)
```

```{r}
inspect(sort(rules, by="lift")[1:10])
```

